---
title: "class 11"
author: "Jenny Zhou"
format: pdf
---

## Identify genetic variants of interest


> Q5: What proportion of the Mexican Ancestry in Los Angeles sample population (MXL) are homozygous for the asthma associated SNP (G|G)?

```{r}
MXL <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(MXL)
```

```{r}
table(MXL$Genotype..forward.strand.)/ nrow(MXL) *100
```

Now let's look at a different population (British in England and Scotland)

```{r}
GBR <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(GBR)
```

```{r}
signif(table(GBR$Genotype..forward.strand.)/ nrow(GBR) *100,3)
```


This variation that is associated with childhood asthma is more frequent the GBR population than the MKL population.

Lets now dig into this further.



## Population Scale Analysis

Determine whether there is any association of the 4 asthma-associated SNPs (rs8067378…) on **ORMDL3** expression.

> Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes.


```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```

```{r}
table(expr$geno)
```
```{r}
exp.med <- function(x) {
  # select one specific genotype
  which.geno <- expr$geno == x
  
  #filter the expression levels for that genotype
  exps <- expr$exp[which.geno]
  #calculate medium of filted expression levels, with 2 decimal points.
  round(median(exps), 2)
}
```

```{r}
exp.med("G/G")
exp.med("A/A")
exp.med("A/G")
```
The median expression level for G/G genotype is 20.07. The median expression level for A/A genotype is 31.25. The median expression level for A/G genotype is 25.06.


> Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the
expression of ORMDL3?

```{r}
library(ggplot2)
```

```{r}
ggplot(expr) + aes(geno, exp, fill = geno) + 
  geom_boxplot(notch=TRUE)
```

SNP effects the expression of ORMDL3. According to the boxplot, comparing to A/A genotype, G/G genotype leads to a lower expression of ORMDL3 by an around 1/3 fold. ORMDL3 expression with a A/G genotype is between that of A/A and G/G.