---
title: "supplement"
author: "Jenny Zhou"
format: html
---
> A. Can you improve this analysis code? 

original code:
```{r}
df <- data.frame(a=1:10, b=seq(200,400,length=10),c=11:20,d=NA) 
df$a <- (df$a - min(df$a)) / (max(df$a) - min(df$a)) 
df$b <- (df$b - min(df$a)) / (max(df$b) - min(df$b)) 
df$c <- (df$c - min(df$c)) / (max(df$c) - min(df$c))
df$d <- (df$d - min(df$d)) / (max(df$a) - min(df$d))

df
```

modified code
```{r}
df <- data.frame(a=1:10, b=seq(200,400,length=10),c=11:20,d=NA)
calculate <- function(x) {
  (x-min(x))/(max(x) - min(x))
}
apply(df, 2,calculate)
```

> Can you improve this analysis code?

original code:
```{r}
library(bio3d) 
s1 <- read.pdb("4AKE") # kinase with drug 
s2 <- read.pdb("1AKE") # kinase no drug 
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA") 
s2.chainA <- trim.pdb(s2, chain="A", elety="CA") 
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b 
s2.b <- s2.chainA$atom$b 
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")

```

> Q1. What type of object is returned from the read.pdb() function? 

> Q2. What does the trim.pdb() function do?

> Q3. What input parameter would turn off the marginal black and grey rectangles in the plots and what do they represent in this case?

> Q4. What would be a better plot to compare across the different proteins?

> Q5. Which proteins are more similar to each other in their B-factor trends. How could you quantify this?





modified code
```{r}
library(bio3d)

ploting <- function(x) {
  read <- read.pdb(x)
  chainA <- trim.pdb(read, chain="A", elety="CA")
  read.b <- chainA$atom$b 
  plotb3(read.b, sse=chainA, typ="l", ylab="Bfactor")
}
```

```{r}
ploting("4AKE")
```


