---
title: "class 10"
author: "Jenny Zhou"
format: pdf
---

# Comparative structure analysis of Adenylate Kinase

adenelate kinase (ADK) is an important drug target and we would love to know how it works.

We will perform PCA on the complete collection of ADK structures in the protein data-bank (PDB)  to reveal detailed features and mechanistic principles of these essential shape changing transitions.


First install the packages needed
> Q10. Which of the packages above is found only on BioConductor and not CRAN? 

msa


> Q11. Which of the above packages is not found on BioConductor or CRAN?: 

Grantlab/bio3d-view

> Q12. True or False? Functions from the devtools package can be used to install packages from GitHub and BitBucket? 

TRUE.


## Search and retrieve ADK structures

Fetch the query sequence for chain A of the PDB ID 1AKE (example ADK sequence). We will use this to find all ADK structures in the PDB.
```{r}
library(bio3d)
aa <- get.seq("1ake_A")
```
```{r}
aa
```

> Q13. How many amino acids are in this sequence, i.e. how long is this sequence? 

214 amino acids.


Now we can use this sequence as a query to BLAST search the PDB to find similar sequences and structures.

```{r}
# Blast or hmmer search 
#b <- blast.pdb(aa)
```

```{r}
# Plot a summary of search results
#hits <- plot(b)
```

Let's see what's in `hits` object.
```{r}
#hits$pdb.id
```
 
 Because we do not want to rerun the code above everytime, we just manually spell the     `hits` vector
```{r}
hits <- NULL
hits$pdb.id <- c('1AKE_A','6S36_A','6RZE_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','3GMT_A','4PZL_A')
```

 

```{r}
# Download releated PDB files
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```


## Align and superpose structures

Align and superpose these structures which are all over place.
```{r}
# Align releated PDBs
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```

Let's have a look at our `pdbs` object.

```{r}
pdbs
```
## Use PCA to analyze the aligned and superposed structure


```{r}
# Perform PCA
pc.xray <- pca(pdbs)
pc.xray
plot(pc.xray)
```

### Color the clusters

rmsd() will calculate all pairwise root-mean-square deviation values of the structural ensemble （distance between different structures）. 
This facilitates clustering analysis based on the pairwise structural deviation:
```{r}
rmsd(pdbs)
```


```{r}
# Calculate RMSD
rd <- rmsd(pdbs)

# Structure-based clustering
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k=3)
grps.rd
```
```{r}
#colored by groups from `cutree`
plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1.1)
```


plot using `ggplot`
```{r}
#Plotting results with ggplot2
library(ggplot2)
library(ggrepel)

# Vector containing PDB codes for figure axis
ids <- basename.pdb(pdbs$id)


df <- data.frame(PC1=pc.xray$z[,1], 
                 PC2=pc.xray$z[,2], 
                 col=as.factor(grps.rd),
                 ids=ids)

p <- ggplot(df) + 
  aes(PC1, PC2, col=col, label=ids) +
  geom_point(size=2) +
  geom_text_repel(max.overlaps = 20) +
  theme(legend.position = "none")
p
```


## Optional further visualization

we can make a wee movie ("trajectory") of the major difference of ADK.
```{r}
# Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```




